DCMIP2016, Part 1: Models and Equation Sets

### 2017-04-22 DCMIP2016-Part1-Models.tex
For FVM: tab, fig and ref updates and corrections

\begin{table}[p]
\caption{Details on the prognostic variables and horizontal discretization for participating dynamical cores.  Equation set indicates whether a model is hydrostatic (H) or non-hydrostatic (NH), and whether the model presently supports the deep-atmosphere formulation (D).  Only two numerical methods are represented among participating models, namely finite-volume (FV) and spectral-element (SE).  More details on horizontal staggering can be found in section \ref{sec:HorizontalStaggering}.} \label{tab:ModelsHorizontal}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Equation set & Prognostic Variables & Horizontal Grid & Numerical & Horizontal  \\
& & & & method & staggering \\ \hline
CSU & NH (Unified) & $\zeta$, $D$, $w$, $p_s$, $\theta_v$, $q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & Z-grid \\
DYNAMICO & H/NH {\color{red}(D)?} & $\vb{v}_h$, $\rho_s w$, $\rho_s$, $\rho_s \theta_v$, $\Phi$, $\rho_s q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & C-grid \\
FV$^3$ & NH & $\vb{u}_h$, $w$, $\rho_s$, $\rho_s \theta_v$, $\Phi$, $\rho_s q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & FV & D-grid \\
FVM & NH (D) & $\rho_d$, $\rho_d \vb{u}_h$, $\rho_d w$, $\rho_d \theta$, ($\Phi$), $\rho_d q$ & Octahedral reduced (\S \ref{sec:grid_oct}) & FV & A-grid \\
GEM & NH & $\vb{u}_h$, $w$, {\color{red}$\log \partial s / \partial \zeta$}, $\log T_v$, $q$ & Yin-yang (\S \ref{sec:grid_yinyang}) & FV & C-grid  \\
HOMME & H & $\vb{u}_h$, $\rho_s$, $T_v$, $\rho_s q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & SE & A-grid \\
ICON & NH & $\vb{u}_h$, $w$, $\rho$, $\theta_v$, $\rho q$ & Icosahedral triangular (\S \ref{sec:grid_ico_tri}) & FV & C-grid \\
MPAS & NH & $\rho_d \vb{u}_h$, $\rho_d w$, $\rho_d$, $\rho_d \theta_v$, $\rho_d q$ & CCVT (\S \ref{sec:grid_ccvt}) & FV & C-grid \\
NICAM & NH & $\rho \vb{u}_h$, $\rho w$, $\rho$, $\rho e$, $\rho q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & A-grid \\
OLAM & NH & $\rho \vb{u}_h$, $\rho w$, $\rho$, $\theta_{il}$, {\color{red} $q$} & Geodesic (\S \ref{sec:grid_geo}) & FV & C-grid \\
Tempest & NH & $\vb{u}_h$, $w$, $\rho$, $\rho \theta_v$, $\rho q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & SE & A-grid \\
\hline 
\end{tabular}
\end{center}
\end{table}

\begin{table}[p]
\caption{Vertical staggering (detailed in section \ref{sec:VerticalStaggering}) and vertical coordinates (detailed in section \ref{sec:VerticalCoordinates}) for participating dynamical cores.} \label{tab:ModelsVertical}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Vertical Staggering & Vertical Coordinate \\ \hline 
CSU & Lorenz & fixed height \\
DYNAMICO & Lorenz & floating mass (\S \ref{sec:FloatingLagrangian}) \\
FV$^3$ & Lorenz & floating mass (\S \ref{sec:FloatingLagrangian}) \\
FVM & Co-located & (floating) height \\
GEM & Modified Charney-Phillips (\S \ref{sec:VerticalStaggering}) & log pressure (\S \ref{sec:GEM_zeta})  \\
HOMME & Co-located & floating mass (\S \ref{sec:FloatingLagrangian}) \\
ICON & Lorenz & fixed height \\
MPAS & Lorenz & fixed height \\
NICAM & Lorenz & fixed height \\
OLAM & Lorenz & fixed height with cut-cells (\S \ref{sec:CutCells}) \\
Tempest & Lorenz & fixed height \\
\hline 
\end{tabular}
\end{center}
\end{table}

\begin{table}[p]
\caption{Standard options for diffusion, stabilization, filters, or fixers in participating dynamical cores (detailed in section \ref{sec:DiffusionStabilization}) and temporal discretization (detailed in section \ref{sec:TemporalDiscretizations}).} \label{tab:ModelsTemporalDiffusion}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Standard options for diffusion, & Temporal discretization \\
& stabilization, filters, or fixers & \\ \hline 
CSU & 4th-order horizontal hyperviscosity & 3rd-order Adams-Bashforth (AB3) \\
DYNAMICO & {\color{red}?} & ARK232 \citep{giraldo2013implicit} \\
FV$^3$ & Divergence damping, monotonic limiting, hyperviscosity & Forward-backwards \citep{LR1997QJR} \\
FVM & Momentum dissipation, scalar diffusion, absorbing layer & Semi-implicit \citep{smolarkiewiczJCP2014} (\S \ref{sec:FVMSemiImplicit}) \\
GEM & Hyperviscosity & Semi-implicit \citep{Girard2014} (\S \ref{sec:GEM_temporal}) \\
HOMME & 4th-order horizontal hyperviscosity & KGU53 \citep{guerra2016high} \\
ICON & Divergence damping, Smagorinsky, hyperdiffusion & Predictor-corrector \\
MPAS & Smagorinsky, hyperdiffusion & Split-explicit \citep{klemp2007conservative} \\
NICAM & 3D divergence damping, Smagornisky, hyperviscosity & Split-explicit \citep{klemp2007conservative} \\
OLAM & Divergence/vorticity damping & 2nd-order Adams-Bashforth, Lax-Wendroff (for tracers) \\
Tempest & 4th-order horizontal hyperviscosity & ARS232 \citep{ascher1997implicit} \\
\hline 
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{table}[p]
\caption{Details on the prognostic variables and horizontal discretization for participating dynamical cores.  Equation set indicates whether a model is hydrostatic (H) or non-hydrostatic (NH), and whether the model presently supports the deep-atmosphere formulation (D).  Only two numerical methods are represented among participating models, namely finite-volume (FV) and spectral-element (SE).  More details on horizontal staggering can be found in section \ref{sec:HorizontalStaggering}.} \label{tab:ModelsHorizontal}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Equation set & Prognostic Variables & Horizontal Grid & Numerical & Horizontal  \\
& & & & method & staggering \\ \hline
CSU & NH (Unified) & $\zeta$, $D$, $w$, $p_s$, $\theta_v$, $q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & Z-grid \\
DYNAMICO & H/NH {\color{red}(D)?} & $\vb{v}_h$, $\rho_s w$, $\rho_s$, $\rho_s \theta_v$, $\Phi$, $\rho_s q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & C-grid \\
FV$^3$ & NH & $\vb{u}_h$, $w$, $\rho_s$, $\rho_s \theta_v$, $\Phi$, $\rho_s q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & FV & D-grid \\
FVM & NH, D & $\rho_d$, $\vb{u}_h$, $w$, $\theta'$, $q$ & Octahedral reduced (\S \ref{sec:grid_oct}) & FV & A-grid \\
GEM & NH & $\vb{u}_h$, $w$, {\color{red}$\log \partial s / \partial \zeta$}, $\log T_v$, $q$ & Yin-yang (\S \ref{sec:grid_yinyang}) & FV & C-grid  \\
HOMME & H & $\vb{u}_h$, $\rho_s$, $T_v$, $\rho_s q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & SE & A-grid \\
ICON & NH & $\vb{u}_h$, $w$, $\rho$, $\theta_v$, $\rho q$ & Icosahedral triangular (\S \ref{sec:grid_ico_tri}) & FV & C-grid \\
MPAS & NH & $\rho_d \vb{u}_h$, $\rho_d w$, $\rho_d$, $\rho_d \theta_v$, $\rho_d q$ & CCVT (\S \ref{sec:grid_ccvt}) & FV & C-grid \\
NICAM & NH & $\rho \vb{u}_h$, $\rho w$, $\rho$, $\rho e$, $\rho q$ & Geodesic (\S \ref{sec:grid_geo}) & FV & A-grid \\
OLAM & NH & $\rho \vb{u}_h$, $\rho w$, $\rho$, $\theta_{il}$, {\color{red} $q$} & Geodesic (\S \ref{sec:grid_geo}) & FV & C-grid \\
Tempest & NH & $\vb{u}_h$, $w$, $\rho$, $\rho \theta_v$, $\rho q$ & Cubed-sphere (\S \ref{sec:grid_cs}) & SE & A-grid \\
\hline 
\end{tabular}
\end{center}
\end{table}

\begin{table}[p]
\caption{Vertical staggering (detailed in section \ref{sec:VerticalStaggering}) and vertical coordinates (detailed in section \ref{sec:VerticalCoordinates}) for participating dynamical cores.} \label{tab:ModelsVertical}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Vertical Staggering & Vertical Coordinate \\ \hline 
CSU & Lorenz & fixed height \\
DYNAMICO & Lorenz & floating mass (\S \ref{sec:FloatingLagrangian}) \\
FV$^3$ & Lorenz & floating mass (\S \ref{sec:FloatingLagrangian}) \\
FVM & Co-located & fixed height \\
GEM & Modified Charney-Phillips (\S \ref{sec:VerticalStaggering}) & log pressure (\S \ref{sec:GEM_zeta})  \\
HOMME & Co-located & floating mass (\S \ref{sec:FloatingLagrangian}) \\
ICON & Lorenz & fixed height \\
MPAS & Lorenz & fixed height \\
NICAM & Lorenz & fixed height \\
OLAM & Lorenz & fixed height with cut-cells (\S \ref{sec:CutCells}) \\
Tempest & Lorenz & fixed height \\
\hline 
\end{tabular}
\end{center}
\end{table}

\begin{table}[p]
\caption{Standard options for diffusion, stabilization, filters, or fixers in participating dynamical cores (detailed in section \ref{sec:DiffusionStabilization}) and temporal discretization (detailed in section \ref{sec:TemporalDiscretizations}).} \label{tab:ModelsTemporalDiffusion}
\begin{center}
\begin{tabular}{cccccc}
\hline Acronym & Standard options for diffusion, & Temporal discretization \\
& stabilization, filters, or fixers & \\ \hline 
CSU & 4th-order horizontal hyperviscosity & 3rd-order Adams-Bashforth (AB3) \\
DYNAMICO & {\color{red}?} & ARK232 \citep{giraldo2013implicit} \\
FV$^3$ & Divergence damping, monotonic limiting, hyperviscosity & Forward-backwards \citep{LR1997QJR} \\
FVM & Monotonic limiting & Semi-implicit \citep{smolarkiewiczJCP2014} (\S \ref{sec:FVMSemiImplicit}) \\
GEM & Hyperviscosity & Semi-implicit \citep{Girard2014} (\S \ref{sec:GEM_temporal}) \\
HOMME & 4th-order horizontal hyperviscosity & KGU53 \citep{guerra2016high} \\
ICON & Divergence damping, Smagorinsky, hyperdiffusion & Predictor-corrector \\
MPAS & Smagorinsky, hyperdiffusion & Split-explicit \citep{klemp2007conservative} \\
NICAM & 3D divergence damping, Smagornisky, hyperviscosity & Split-explicit \citep{klemp2007conservative} \\
OLAM & Divergence/vorticity damping & 2nd-order Adams-Bashforth, Lax-Wendroff (for tracers) \\
Tempest & 4th-order horizontal hyperviscosity & ARS232 \citep{ascher1997implicit} \\
\hline 
\end{tabular}
\end{center}
\end{table}
